---
title: "Kaplan-Meier Exploration"
output: html_document
---

# Kaplan-Meier Survival Models

Playing around with Kaplain-Meier estimation of survival curve.

 * KM produces a basic survival curve solely using the data available (non-parametric)
 * KM is equivalent to an actuarial approach to developing mortality rates using a cohort.
   * See section 4.2 from the [SOA's Experience Study Calculations](https://www.soa.org/globalassets/assets/Files/Research/2016-10-experience-study-calculations.pdf) paper.
   * In the example in section 4.2, withdrawals are equivalent to censoring in survival analysis, and deaths are equivalent to the event being studied.

```{r include=FALSE}
#install.packages(c("survival", "survminer", "ranger", "Rcpp", "ggfortify))

library("survival")
library("survminer")
library("tidyverse")
```


Get the test dataset.  Target is week and arrest, week is the length of time, and arrest is the censoring variable.  8 predictor variables are single columns, while one variable is split out over all the 52 weeks (emp1-52).

Reference for the dataset:
https://www.rdocumentation.org/packages/RcmdrPlugin.survival/versions/1.2-1/topics/Rossi

```{r}
url <- "http://socserv.mcmaster.ca/jfox/Books/Companion/data/Rossi.txt"
rossi <- read.table(url, header=TRUE)
print(paste0(c("Rows: ", "Cols: "), dim(rossi)))
print(names(rossi))
#summary(rossi[])
```


Histogram of when people either get arrested or are lost from the study.
```{r}
hist(rossi$week)
```

Fit Kaplan-Meier with `survfit` and use a survival object as the target in the formula using `Surv(time, censor_var)`.
```{r}
km_rossi <- survfit(Surv(week, arrest) ~ 1, data = rossi)
plot(km_rossi, ylim=c(0.7, 1), xlab="Weeks", ylab="Proportion Not Rearrested")
```

Access the estimated survival function using `$surv`.  Calculations below show the hazard at each time.  Similar to a life insurance q from a mortality table.
```{r}
surv_calcs <- tibble("time" = km_rossi$time,
                     "surv" = km_rossi$surv,
                     "surv_lag" = dplyr::lag(km_rossi$surv, default = 1))

surv_calcs <- surv_calcs %>%
  mutate(haz = 1 - (surv / surv_lag))

ggplot(surv_calcs) + geom_line(aes(x=time, y=haz))
```

Change the formula to have the Kaplan-Meier plot split by the variable `fin`. It now produces two separate survival curves from the data.
```{r}
survfit(Surv(week, arrest) ~ fin, data = rossi) %>%
plot(., ylim=c(0.7, 1))

```

`survminer::ggsurvplot` gives a slightly nicer looking output plot for the survival curves.  Here are the resulting curves when the data is partitioned into 4 subsets defined by `fin` and `mar`.

```{r, warning=FALSE}
survfit(Surv(week, arrest) ~ fin + mar, data = rossi) %>%
ggsurvplot(., ylim=c(0.6, 1))

```

